Impact of grids and dynamical cores in CESM2.2 on the meteorology and climate of the Arctic

### 2021-12-23 arcticgrids.tex
writing

All grids and dycores are run using an identical transient 1979-1998 AMIP-style configuration, with prescribed monthly SST/sea-ice after \cite{CESMSST}. This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2. The Community Land Model (CLM5; \url{https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/index.html}), which uses the same grid as the atmosphere grid, calculates the surface energy balance at each land tile within a grid cell, which is used to compute snow and bare ice melting to inform the surface mass balance (SMB) of a glacier unit \cite{VKETAL2020JGR}. For land ice tiles coinciding with the Antarctic and Greenland Ice Sheets, the surface energy and mass balance are evaluated at ten elevation classes, and integrated over sub-grid hypsometric area-elevation bins for more accurate estimation of the surface mass balance of the big ice sheets. Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation. Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up  by forcing CLM5 offline, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years. The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.

@@

Comparing the variable-resolution grids to the uniform resolution grids is complicated because we are simultaneously increase the grid resolution and reducing the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}. An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$. Figure~\ref{fig:dT-hires} shows the change in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$. A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB. Figure~\ref{fig:dT-hires} shows the difference in climatological temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid. The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.

@@

As has been previously reported, horizontal resolution can have large impacts on extreme precipitation events. Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with synoptic storms, by month. The $ARCTIC$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year. This is primarily due to an increase in resolution; the $ne30pg3^{*}$ only marginally increases the precipitation rates compared with $ne30pg3$. It should also be noted that the PDFs are evaluated on an identical composite grid, and so storm statistics are not impacted by differences in dataset resolution.

@@

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=60mm]{figs/temp_grisres_dxoverglc.pdf}&
         \includegraphics[width=60mm]{figs/temp_grisres_error.pdf} \\
\end{tabular}
\end{center}
\caption{.}
\label{fig:prect}
\end{figure}

>>>>

All grids and dycores are run using an identical transient 1979-1998 AMIP-style configuration, with prescribed monthly SST/sea-ice after \cite{CESMSST}. This configuration refers to the $FHIST \ compset$ and runs out of the box in CESM2.2. The Community Land Model (CLM5; \url{https://escomp.github.io/ctsm-docs/versions/release-clm5.0/html/users_guide/index.html}), which uses the same grid as the atmosphere grid, calculates the surface energy balance at each land tile within a grid cell, which is used to compute snow and bare ice melting to inform the surface mass balance (SMB) of a glacier unit \cite{VKETAL2020JGR}. For land ice tiles coinciding with the Antarctic and Greenland Ice Sheets, the surface energy and mass balance are evaluated at ten elevation classes, and integrated over sub-grid hypsometric area-elevation bins for more accurate estimation of the surface mass balance of the big ice sheets. Ice accumulation is modeled as a capping flux, or snow in excess of the assumed 10 m snow cap, and refreezing of liquid within the snowpack additionally acts as a source of mass in the SMB calculation. Since the 10 m snowcap needs to be reached in the accumulation zone to simulate the SMB, the snow depths in the variable-resolution grids were spun-up  by forcing CLM5 standalone, cycling over 20 years of a fully coupled $ARCTIC$ run for about 500 years. The uniform resolution grids are all initialized with an SMB from an existing $f09$ spun-up initial condition.

\subsection{SMB Analysis}

A common high resolution dataset is used to generate the GrIS boundary conditions using the CLM dataset creation tools. Since we are interested in the total ice sheet SMB, we seek to integrate various components of the SMB over a common ice mask to get the total mass change over the GrIS. Figure~\ref{fig:grisdx} shows the GrIS ice mask area across the different grids, as a function of the number of grid points. Due to the use conservative regridding in the CLM tools, the interpolation errors are small and ice mask areas have less than $1.5\%$ errors relative to the raw ice mask dataset. RACMO, however, uses a smaller ice mask, about $3\%$ smaller than the raw CLM dataset. While Figure~\ref{fig:grisdx} suggests integrating quantities over the native ice mask of the six grids would probably not suffer from large errors due to differing ice masks, we seek to compare these integrated quantities with RACMO. Therefore, we have taken the approach of mapping all model fields to the lowest resolution grids and integrating over the respective low resolution ice masks. Due to the sensitivity of mapping errors to grid coordinates (i.e., unstructured or structured), all quantities are evaluated on both the $f19$ and $ne30pg2$ low resolution grids. In addition, two remapping packages are used; ESMF conservative and a TempestRemap high-order, monotone algorithm. In all, each integrated quantity is evaluated (at most) four times to provide an estimate of uncertainty due to differences in grid coordinates and remapping algorithm.

@@

Comparing the variable-resolution grids to the uniform resolution grids is complicated because we simultaneously increase the grid resolution and reduce the physics time-step, both which noticeably impact the solution \cite{W2008TELLUS}. An additional $ne30pg3$ simulations is run with the physics time-step used in the $ARCTIC$ grid, referred to as $ne30pg3^{*}$. Figure~\ref{fig:dT-hires} shows the change in climatological temperature in zonal-mean height space between $ne30pg3^{*}$ and $ne30pg3$. A similar warming response to increasing resolution occurs when the time-step is reduced, and the mechanism is similar in that the shorter time-step facilitates greater condensational heating by CLUBB. Figure~\ref{fig:dT-hires} shows the difference in climatological temperature between the $ARCTIC$ grid and the $ne30pg3^{*}$ grid. The greater condensational heating and warmer temperatures are confined to the regionally refined region when the impact of physics time-steps is removed from the analysis.

@@

As has been previously reported, horizontal resolution can have large impacts on extreme precipitation events. Figure~\ref{fig:comp-pdf} is a PDF of the precipitation rates associated with synoptic storms, by month. The $ARCTIC$ run has larger extreme precipitation rates compared to $ne30pg3$ in every month, but the increase is greatest in the summer months, which coincides with the most extreme events of the year. This is primarily due to an increase in resolution; the $ne30pg3^{*}$ only marginally increases the precipitation rates compared with $ne30pg3$. The $f09$ and $f19$ grids looks similar to $ne30pg3$, but with $f09$ having more frequent extreme precipitation events than $f19$ (not shown). It should also be noted that the PDFs are evaluated on an identical composite grid for all runs, and so storm statistics are not impacted by differences in output resolution.

@@

The surface mass balance (SMB) of the Greenland Ice Sheet (GrIS) is simulated in all grids and dycores in this study. As the accuracy of the computed SMB is sensitive to grid resolution, Figure~\ref{fig:grisdx} shows the average grid spacing for all six grids in this study. The $ne30pg2$ grid has the coarsest representation with an average $\Delta x=160~km$, and the $ARCTICGRIS$ grid has the highest resolution with an average $\Delta x=14.6~km$, similar to the grid spacing of the $11~km$ RACMO model. The $ne30pg3$ grid has an average $\Delta x=111.2~km$, which is substantially more coarse than the $f09$ grid, having an average $\Delta x=60~km$. This is interesting because $ne30pg3$ and $f09$ have similar average grid spacing over the entire globe, and comparable computational costs, but due to the convergence of meridians the finite-volume model has enhanced resolution over GrIS. The $ARCTIC$ grid has an average grid spacing over GrIS of $\Delta x=27.8~km$, and is about 10 times more expensive than the $1^{\circ}$ models. The $ARCTICGRIS$ grid is about twice as expensive as the $ARCTIC$ grid.

@@

\begin{figure}[t]
\begin{center}
\begin{tabular}{cccc}
         \includegraphics[width=60mm]{figs/temp_grisres_dxoverglc.pdf}&
         \includegraphics[width=60mm]{figs/temp_grisres_error.pdf} \\
\end{tabular}
\end{center}
\caption{.}
\label{fig:grisdx}
\end{figure}
